---
title: "Example code for R S3 object oriented programming"
author: "Kevin Rue-Albrecht"
date: "16/11/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Exercises

## Character

```{r}
character_object <- "A"
character_object
```

```{r}
str(character_object) # structure of object
```

```{r}
class(character_object)
```

```{r}
attributes(character_object) # eg if the vector had names you would see this here
```

```{r}
character_object2 <- c("A", "B", "C")
character_object2
```

```{r}
str(character_object2)
```

```{r}
class(character_object2)
```

```{r}
attributes(character_object2)
```

# Exercise

## Factors

```{r, include=TRUE}
factor_object <- factor(c("A", "B", "C")) # A factor has levels. All values in factor must take one of these levels
factor_object
```

```{r}
str(factor_object) # look up table - each value of factor assigned a number 
```

```{r}
class(factor_object)
```

```{r}
attributes(factor_object)
```

```{r}
attr(factor_object, "levels")
```

```{r}
attr(factor_object, "class")
```

# Exercise

## Data-frames
NB a dataframe is a list where the column lengths must be equal so that the dataframe can presrnt as a coherent table

```{r}
dataframe_object <- data.frame(
  numbers = c(1, 2, 3),
  strings = c("A", "B", "C")
  )
dataframe_object
```

```{r}
class(dataframe_object)
```

```{r, include=TRUE}
str(dataframe_object)
```

```{r}
attributes(dataframe_object)
```

# Exercise

## Matrices
NB a matrix is a vector, shaped across a number of rows and columns
naturally fills by columns; can parse an argument to flip this around and fill by rows instead
All values in matrix must be of the same type 
```{r}
matrix_object <- matrix(
  data = c(1, 2, 3, 4, 5, 6, 7, 8),
  nrow = 2
  )
matrix_object
```

```{r}
class(matrix_object)
```

```{r}
str(matrix_object)
```

```{r}
attributes(matrix_object)
```

```{r}
attr(matrix_object, "dim")
```

# Exercise

## S3 method dispatch
! Never include dots in naming in R: instead use _ letters and digits
```{r}
methods("print") %>% head(14)
```

```{r}
print.factor(factor_object)
```

```{r}
print.factor(character_object)
```

```{r}
print.default(character_object)
```

